equations

### 2023-05-29 Manuscript1.Rmd
cleaning

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, osfr, lmerTest, MuMIn, DHARMa, here, cmdstanr, tidybayes, bayesplot, posterior,future,furrr, patchwork, R.utils, brms, extraDistr,cowplot)

@@

## R Markdown

```{r parameters visualization, fig.height=14, fig.width=16}
dd = get_experiment()

@@

layout = (c(
  area(1, 1, 1,1),
  area(1, 2, 1,2),
  area(1, 3, 1,3)

@@

p1 = kalmanplot[[1]]+rwplot[[1]]+wbplot[[1]]+plot_layout(design = layout)+plot_annotation(tag_levels = list(c('B', 'C',"D","E","F","G")))
p2 = kalmanplot[[2]]+rwplot[[2]]+wbplot[[2]]+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("E","F","G")))

p1
p2

RW_logistic = RW_logistic_vis(dd)
```

```{r parameter recovery for RW}
mu_alpha = seq(0.1,0.5,by = 0.1)
kappa_alpha = seq(10,300,length.out = 3)

@@

rw_pp = furrr::future_map(data_list, ~parameter_recovery_rw(.x), .progress = TRUE, .options = furrr_options(seed = 123))

@@

nsubs = 50

@@

```{r}

files = paste0(here::here("workspace","parameterrecovery"),"/",list.files(here::here("workspace","parameterrecovery")))

#plot parameterrecovery
load(files[[2]])
q = lapply(here("scripts", scripts), source)
rwplot = plot_parameterrecovery(rw_pp, div = T, subs = 5)
load(files[[3]])
q = lapply(here("scripts", scripts), source)
wbplot = plot_parameterrecovery(wb_pp, div = T, subs = 5)
load(files[[1]])
q = lapply(here("scripts", scripts), source)
kalmanplot = plot_parameterrecovery(kalman_pp, div = T, subs = 5)

@@

plot(layout)

((rwplot[[1]]+wbplot[[1]]+kalmanplot[[1]])/(rwplot[[2]]+wbplot[[2]]+kalmanplot[[2]])/((rwplot[[3]]+wbplot[[3]]+kalmanplot[[3]])))+
  plot_annotation(tag_levels = 'A')+
  plot_layout(design = layout,guides = 'collect')

@@

```{r model recovery}
########################################### Resorla wagner parameters:
mu_alpha = rep(0.3,100)
kappa_alpha = 50
mu_w1 = rep(0.3,100)
kappa_w1 = 50
sd_precision_percept = 0.5
sd_beta = 0.5
nsubs = 10
parameters = expand.grid(mu_alpha = mu_alpha,
                         kappa_alpha = kappa_alpha,
                         mu_w1 = mu_w1,
                         kappa_w1 = kappa_w1,
                         sd_precision_percept = sd_precision_percept,
                         sd_beta = sd_beta,
                         nsubs = nsubs)

@@

mu_b2 = 2
sd_b2 = 1

@@

@@

data_list = data_list_logs

@@

```{r model recovery table:}

@@

get_value(rw_loo)
```

@@

```{r plotting 1 agent}

percept = sum %>% filter(str_detect(variable, "painMu"))
percept = percept[1:160,]

expect = sum %>% filter(str_detect(variable, "expectMu"))
expect = expect[1:160,]

percept$stim = data$stim
percept$expect = expect$mean

percept %>% ggplot(aes(x=1:nrow(.), y = mean, col = as.factor(stim)))+geom_point()

percept %>% ggplot(aes(x=expect, y = mean, col = as.factor(stim)))+geom_point()

ass = sum %>% filter(str_detect(variable, "associ"))
ass = ass[1:161,]

ass %>% ggplot(aes(x=1:nrow(.), y = mean))+geom_line()+geom_point(data = data, aes(x = trial, y = u))

rw_fit_sess1$diagnostic_summary()

alpha = get_diag(rw_fit_sess1, "alpha")
w1 = get_diag(rw_fit_sess1, "w1")
pp = get_diag(rw_fit_sess1, "precision_percept")
beta = get_diag(rw_fit_sess1, "beta")

@@

@@

@@

```{r}
get_corplot(rw_fit_sess1, rw_fit_sess2, "alpha")
get_corplot(rw_fit_sess1, rw_fit_sess2, "w1")
get_corplot(rw_fit_sess1, rw_fit_sess2, "precision_percept")

get_corplot(wb_fit_sess1, wb_fit_sess2, "alpha")
get_corplot(wb_fit_sess1, wb_fit_sess2, "w1")
get_corplot(wb_fit_sess1, wb_fit_sess2, "w2")
get_corplot(wb_fit_sess1, wb_fit_sess2, "beta")
get_corplot(wb_fit_sess1, wb_fit_sess2, "percept_precision")

```{r}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0)

kalman_fit_sess1 <- mod$sample(
      data = data1,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )
```

```{r hierachical session correlation (RW)}

fit <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 100,
  adapt_delta = 0.90,
  max_treedepth = 12
)

sum = fit$summary(variables = c("mu_alpha","mu_w1","sd_beta","sd_precision_percept", "kappa_alpha","kappa_w1","alpha","w1","beta","precision_percept"))

variabless = c("mu_alpha[1]","mu_w1[1]","sd_beta[1]","sd_precision_percept[1]","mu_alpha[2]","mu_w1[2]","sd_beta[2]","sd_precision_percept[2]")

pairs = mcmc_pairs(fit$draws(variables = variabless), np = nuts_params(fit), pars = variabless,
                   off_diag_args = list(size = 0.75))

mcmc_trace(fit$draws(variables = "alpha"))

alpha = fit$summary(variables = "alpha")
w1 = fit$summary(variables = "w1")

alphas1 <- w1 %>%
  filter(grepl("1]$", variable))

alphas2 <- w1 %>%
  filter(grepl("2]$", variable))

cor.test(alphas1$mean, alphas2$mean)

correlation = as_draws_df(fit$draws(variables = "correlation_alpha"))
hist(correlation$correlation_alpha)
correlation = as_draws_df(fit$draws(variables = "correlation_w1"))
hist(correlation$correlation_w1)
```

fit_wb_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 100,
  adapt_delta = 0.90,
  max_treedepth = 12
)

```{r fitting to both sessions to hierachiacal RW}

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(session == 1) %>% filter(id %in% unique(id)[1])),
            sessions = 2,
            pred = get_3darray(data,variable = "pred"),
            u = get_3darray(data,variable = "u")
            )

mod = cmdstan_model(here::here("Stan","rw_vs_rw_hier.stan"))

rw_fit_ses1 <- mod$sample(
      data = data1,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

sum = rw_fit_ses1$summary(variables = c("alpha", "mu_alpha","sd_alpha","mu_mu_alpha","sd_sd_alpha","mu_sd_alpha","sd_mu_alpha"))
```

```{r, fitting to real data tester hierachical RW sessions1}

rw_fit_ses1 <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

sum1 = rw_fit_ses1$summary()
```

```{r, fitting to real data tester hierachical RW sessions2}

rw_fit_ses2 <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

sum2 = rw_fit_ses2$summary()

get_corplot(rw_fit_ses1,rw_fit_ses2, "b2_b")
```

```{r}

fit <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 10,
  adapt_delta = 0.90,
  max_treedepth = 12
)

```{r}
mu_alpha = 0.3
kappa_alpha = 30

mu_b0 = -5
sd_b0 = 1

mu_b1 = -1
sd_b1 = 1

mu_b2 = 1.3
sd_b2 = 1.3

mu_b0_b = -5
sd_b0_b = 2

mu_b1_b = 0.5
sd_b1_b = 0.5

mu_b2_b = 0.5
sd_b2_b = 0.5

df = hier_rw_logistic(parameters)

data = df[[1]]

hier = df[[2]]
df = df[[1]]

fit <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

hier_parameters = c("mu_b0","mu_b1","mu_b2","sd_b0","sd_b1","sd_b2","mu_b0_b","mu_b1_b","mu_b2_b","sd_b0_b","sd_b1_b","sd_b2_b","kappa_alpha","mu_alpha","sd_percept_precision")

sub_parameters = c("b0","b1","b2","b0_b","b1_b","b2_b","alpha","percept_precision")

list(hier = data.frame(fit$summary(hier_parameters),
                                  reals = hier %>% dplyr::select(all_of(hier_parameters)) %>% pivot_longer(everything()) %>% rename(reals = value),
                                  div = sum(fit$diagnostic_summary()$num_divergent),index = index),
                sub = data.frame(fit$summary(sub_parameters) %>% arrange(variable),
                                 reals = df %>% filter(trial == 1) %>% mutate(ids = 1:nrow(.)) %>% dplyr::select(all_of(sub_parameters),ids) %>% 
                                 pivot_longer(cols = -ids) %>% rename(reals =value) %>% mutate(names = paste0(name,"[",ids,"]")) %>% arrange(names) %>% select(names,reals),
                                 div = sum(fit$diagnostic_summary()$num_divergent),index = index))

#sum = fit$summary()

>>>>

### loading packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, osfr, lmerTest, MuMIn, DHARMa, here, cmdstanr, tidybayes, bayesplot, posterior,future,furrr, patchwork, R.utils, brms, extraDistr,cowplot,HDInterval)

@@

### parameter visualization for the four models (figure 1 & 2)
```{r parameter visualization, fig.height=10, fig.width=12}
dd = get_experiment()

@@

RW_logistic = RW_logistic_vis(dd)

layout = (c(
  area(1, 1, 1,1),
  area(1, 2, 1,2)

@@

p1 = RW_logistic[[1]]+rwplot[[1]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c('B', 'C')))
p2 = wbplot[[1]]+kalmanplot[[1]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("D","E")))
p3 = RW_logistic[[2]]+rwplot[[2]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("A","B")))
p4  = wbplot[[2]]+kalmanplot[[2]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("C","D")))

p1
p2
p3
p4

### parameter recovery for the four models
```{r parameter recovery for RW}
mu_alpha = seq(0.1,0.5,by = 0.1)
kappa_alpha = seq(10,300,length.out = 3)

@@

rw_pp = furrr::future_map(data_list, ~parameter_recovery_rw(.x), .progress = TRUE, .options = furrr_options(seed = 123))
```

@@

@@

#plotting parameter recovery (figure 3)
```{r}
files = paste0(here::here("workspace","parameterrecovery"),"/",list.files(here::here("workspace","parameterrecovery")))

#plot parameterrecovery
load(files[[1]])
q = lapply(here("scripts", scripts), source)
kalmanplot = plot_parameterrecovery(kalman_pp, div = T, subs = 5)
load(files[[2]])
q = lapply(here("scripts", scripts), source)
logisticplot = plot_parameterrecovery(rw_pp_logs, div = F, subs = 5)
load(files[[3]])
q = lapply(here("scripts", scripts), source)
rwplot = plot_parameterrecovery(rw_pp, div = T, subs = 5)
load(files[[4]])
q = lapply(here("scripts", scripts), source)
wbplot = plot_parameterrecovery(wb_pp, div = T, subs = 5)

@@

((rwplot[[1]]+wbplot[[1]])/(kalmanplot[[1]]+logisticplot[[1]]))+
  plot_annotation(tag_levels = 'A')+
  plot_layout(guides = 'collect')

((rwplot[[2]]+wbplot[[2]])/(kalmanplot[[2]]+logisticplot[[2]]))+
  plot_annotation(tag_levels = 'A')+
  plot_layout(guides = 'collect')

@@

### model recovery for the three computation models
```{r model recovery}
########################################### Resorla wagner parameters:
mu_alpha = rep(0.3,100)
kappa_alpha = 50
mu_w1 = 0.3
kappa_w1 = 50
sd_precision_percept = 0.5
sd_beta = 0.5
nsubs = 10
parameters = expand.grid(mu_alpha = mu_alpha,
                         kappa_alpha = kappa_alpha,
                         mu_w1 = mu_w1,
                         kappa_w1 = kappa_w1,
                         sd_precision_percept = sd_precision_percept,
                         sd_beta = sd_beta,
                         nsubs = nsubs)

@@

mu_b2 = 1
sd_b2 = 1

mu_b0_b = -3
sd_b0_b = 0.5

mu_b1_b = 5
sd_b1_b = 0.5

@@

plan(multisession, workers = 10)

@@

data_list = data_list_logs[1:50]

plan(multisession, workers = 10)

@@

### table 1
```{r model recovery table}

files = paste0(here::here("workspace","model recovery"),"/",list.files(here::here("workspace","model recovery")))

#plot parameterrecovery
load(files[[1]])

@@

kalmandata = get_value(kalman_loo)
rwdata = get_value(rw_loo)
wbdata = get_value(wb_loo)

### fitting the models to session 1's data

@@

@@

```{r, fitting to real data tester hierachical logs RW sessions1}

rw_logs_fit_ses1 <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

sum1 = rw_logs_fit_ses1$summary()
```

### Model comparison on sessions 1 (table 2)
```{r, model comparison}
#rw = rw_fit_sess1$loo()
#wb = wb_fit_sess1$loo()
#rw_logs = rw_logs_fit_ses1$loo()

rw = as_draws_df(rw_fit_sess1$draws(variables = "log_lik"))
wb = as_draws_df(wb_fit_sess1$draws(variables = "log_lik"))
rw_logs = as_draws_df(rw_logs_fit_ses1$draws(variables = "log_lik"))

rows_with_na <- which(apply(is.na(wb), 1, any))
wb = wb[-rows_with_na,]
rw = rw[-rows_with_na,]
rw_logs = rw_logs[-rows_with_na, ]

wb = loo::loo(as.matrix(wb))
rw = loo::loo(as.matrix(rw))
rw_logs = loo::loo(as.matrix(rw_logs))

loo_compare(rw,wb,rw_logs)
loo_model_weights(list(rw,wb,rw_logs))
```

### fitting the models to session 2's data

@@

### fitting the models to session 2's data

wb_fit_sess2$diagnostic_summary()

### fitting the models to session 2's data

```{r, fitting to real data tester hierachical logs RW sessions2}

rw_logs_fit_ses2 <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

### Model comparison on sessions 2 (table 2)

```{r, model comparison}
#rw = rw_fit_sess1$loo()
#wb = wb_fit_sess1$loo()
#rw_logs = rw_logs_fit_ses1$loo()

rw = as_draws_df(rw_fit_sess2$draws(variables = "log_lik"))
wb = as_draws_df(wb_fit_sess2$draws(variables = "log_lik"))
rw_logs = as_draws_df(rw_logs_fit_ses2$draws(variables = "log_lik"))

rows_with_na <- which(apply(is.na(wb), 1, any))
wb = wb[-rows_with_na,]
rw = rw[-rows_with_na,]
rw_logs = rw_logs[-rows_with_na, ]

wb = loo::loo(as.matrix(wb))
rw = loo::loo(as.matrix(rw))
rw_logs = loo::loo(as.matrix(rw_logs))

loo_compare(rw_logs,rw,wb)

loo_model_weights(list(rw_logs,rw,wb))
```

### Fitting the models to both sessions

```{r hierachical session correlation (RW)}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0) %>% mutate(id = as.factor(id), trial = as.factor(trial), session = as.factor(session))

fit_rw_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 100,
  adapt_delta = 0.90,
  max_treedepth = 12
)

alpha_wb = get_corplotreal(fit_rw_bs, "alpha", ymin = 0.1, ymax = 0.2, xmin = 0.2, xmax = 0.3)
w1_wb = get_corplotreal(fit_rw_bs, "w1", ymin = 0.125, ymax = 0.3, xmin = 0.6, xmax = 0.75)

HDInterval::hdi(as_draws_df(fit_rw_bs$draws(variables = "correlation_alpha")) %>% select(starts_with("correlation")))

HDInterval::hdi(as_draws_df(fit_rw_bs$draws(variables = "correlation_w1")) %>% select(starts_with("correlation")))

fit_wb_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 50,
  adapt_delta = 0.90,
  max_treedepth = 12
)

fit_wb_bs$diagnostic_summary()

alphaplot1 = get_corplot(wb_fit_sess1, wb_fit_sess2, "alpha")

alpha_wb = get_corplotreal(fit_wb_bs, "alpha", ymin = -0.075, ymax = 0.325, xmin = 0.65, xmax = 1)
w1_wb = get_corplotreal(fit_wb_bs, "w1", ymin = 0.02, ymax = 0.12, xmin = 0.19, xmax = 0.27)
w2_wb = get_corplotreal(fit_wb_bs, "w2", ymin = 0.1, ymax = 0.2, xmin = 0.4, xmax = 0.6)

HDInterval::hdi(as_draws_df(fit_wb_bs$draws(variables = "correlation_alpha")) %>% select(starts_with("correlation")))

HDInterval::hdi(as_draws_df(fit_wb_bs$draws(variables = "correlation_w1")) %>% select(starts_with("correlation")))

HDInterval::hdi(as_draws_df(fit_wb_bs$draws(variables = "correlation_w2")) %>% select(starts_with("correlation")))

```{r logistic both sessions}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0) %>% mutate(id = as.factor(id), trial = as.factor(trial), session = as.factor(session))

fit_rw_logs_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 50,
  adapt_delta = 0.90,
  max_treedepth = 12
)

HDInterval::hdi(get_correlation(rw_logs_fit_ses1,rw_logs_fit_ses2,"alpha"))

HDInterval::hdi(get_correlation(rw_logs_fit_ses1,rw_logs_fit_ses2,"b1"))

HDInterval::hdi(get_correlation(rw_logs_fit_ses1,rw_logs_fit_ses2,"b2"))

### plotting the correlation (figure 4)

```{r correlation plots}

#logs

alpha_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"alpha",ymin = 0, ymax = 0.3, xmin = 0.55, xmax = 0.85)
#b0_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"b0",ymin = -4.25, ymax = -3, xmin = -3, xmax = -2)
b1_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"b1",ymin = 0.2, ymax = 1.5, xmin = 2.5, xmax = 3.4)
b2_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"b2",ymin = -0.2, ymax = 0, xmin = 0.3, xmax = 0.575)

rw_logs = alpha_logs+b1_logs+b2_logs
#wb
alpha_wb = get_corplotreal(fit_wb_bs, "alpha", ymin = -0.075, ymax = 0.275, xmin = 0.56, xmax = 0.91)
w1_wb = get_corplotreal(fit_wb_bs, "w1", ymin = 0.02, ymax = 0.1, xmin = 0.17, xmax = 0.25)
w2_wb = get_corplotreal(fit_wb_bs, "w2", ymin = 0.1, ymax = 0.21, xmin = 0.3, xmax = 0.55)

wb = alpha_wb+w1_wb+w2_wb
#rw
alpha_rw = get_corplotreal(fit_rw_bs, "alpha", ymin = 0.1, ymax = 0.2, xmin = 0.2, xmax = 0.3)
w1_rw = get_corplotreal(fit_rw_bs, "w1", ymin = 0.125, ymax = 0.3, xmin = 0.6, xmax = 0.75)

rw = alpha_rw+w1_rw

rw+plot_annotation(tag_levels = list(c("A",'B')))
wb+plot_annotation(tag_levels = list(c("C",'D',"E")))
rw_logs+plot_annotation(tag_levels = list(c("F",'G',"H")))
```

### fitting the kalman model just to see

@@

### 2023-05-29 equations.Rmd
updates
cleaning

---
title: "equations"
author: "Jesper Fischer Ehmsen"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

@@

>>>>

---
title: "equations"
author: "Jesper Fischer Ehmsen"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

@@

#supplementary

showing equation (15 and 16)
Starting of with the mean of the precept

$$
\begin{equation}
\mu_{\psi_t} = \frac{\sigma_{s}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}*\mu_{E_{t-1}}+\frac{\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}*S_t
\end{equation}
$$
using (13) we need to show the following

$$
\begin{equation}
(1-\gamma) = \frac{\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
$$
Now uterlising the fact that if $x = \frac{a}{b}$ then $1-x = \frac{(b-a)}{b}$

$$
\begin{equation}
(1-\gamma) = \frac{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2-\sigma_{s}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
$$
which gives exactly

$$
\begin{equation}
(1-\gamma) = \frac{\sigma_{\psi}^2+\sigma_{E_t}^2}{\sigma_{s}^2+\sigma_{\psi}^2+\sigma_{E_t}^2}
\end{equation}
$$
